---
title: "DataTrans_DatTidy_Lab"
author: "Na Zhou"
date: "Dec 2nd 2024"
output: html_document
---

###Data Transformation

##Start by loading library(tidyverse) and library(dplyr) if not done already.

```{r}
library(tidyverse)
library(dplyr)
library(nycflights13)
View(flights)
```


3.2.5 Exercises
In a single pipeline for each condition, find all flights that meet the condition:

1. Had an arrival delay of two or more hours
2. Flew to Houston (IAH or HOU)
3. Were operated by United, American, or Delta
4. Departed in summer (July, August, and September)
5. Arrived more than two hours late but didn’t leave late
6. Were delayed by at least an hour, but made up over 30 minutes in flight
7. Sort flights to find the flights with the longest departure delays. Find the flights that left earliest in the morning.

8. Sort flights to find the fastest flights. (Hint: Try including a math calculation inside of your function.)

9. Was there a flight on every day of 2013?

10. Which flights traveled the farthest distance? Which traveled the least distance?

11. Does it matter what order you used filter() and arrange() if you’re using both? Why/why not? Think about the results and how much work the functions would have to do.

```{r}
#1. Had an arrival delay of two or more hours
answer_1 <- flights |> 
              filter(arr_delay >= 120)
answer_1
#2. Flew to Houston (IAH or HOU)
answer_2 <- flights |>
              filter(dest %in% c("IAH", "HOU"))
answer_2
#3. Were operated by United, American, or Delta
answer_3 <- flights |>
              filter(carrier %in% c("UA", "AA", "DL"))
answer_3
#4. Departed in summer (July, August, and September)
answer_4 <- flights |>
              filter(month %in% c(7, 8, 9))
answer_4
#5. Arrived more than two hours late but didn’t leave late
answer_5 <- flights |> 
              filter(arr_delay > 120, dep_delay <= 0)
answer_5
#6. Were delayed by at least an hour, but made up over 30 minutes in flight
answer_6 <- flights |>
              filter(dep_delay >= 60, (dep_delay - arr_delay) > 30)
answer_6
#7. Sort flights to find the flights with the longest departure delays. 
answer_7_1 <- flights |>
                arrange(desc(dep_delay)) |> 
                slice(1)
answer_7_1
#Find the flights that left earliest in the morning.
answer_7_2 <- flights |>
                arrange(dep_time) |> 
                slice(1)
answer_7_2
#8. Sort flights to find the fastest flights. (Hint: Try including a math calculation inside of your function.)
answer_8 <- flights |>
              mutate(speed = distance / (air_time / 60)) |> 
              arrange(desc(speed)) |>
              relocate(speed) |> 
              slice(1)
answer_8
#9. Was there a flight on every day of 2013? Yes, there was a flight on every day of 2013
answer_9 <- flights |>
              distinct(year, month, day) |> 
              nrow()
answer_9
#10. Which flights traveled the farthest distance? 
answer_10_1 <- flights |>
                  arrange(desc(distance)) |> 
                  slice(1)
answer_10_1
#Which traveled the least distance?
answer_10_2 <- flights |>
                  arrange(distance) |> 
                  slice(1)
answer_10_2
#11. Does it matter what order you used filter() and arrange() if you’re using both? Why/why not? Think about the results and how much work the functions would have to do.
#Yes, it matters the order I use filter() and arrange(), if I use filter() first, it can help me to reduce rows to arrange, if I use arrange() first, it will sort all rows first which causes more computation efficiency and time and I don't need all rows to be sorted either
```



###Displaying three tables

TB cases in world

```{r}
table1
```
```{r}
table2
```
```{r}
table3
```
```{r}
# Compute rate per 10,000
table1 |>
  mutate(rate = cases / population * 10000)


# Compute total cases per year
table1 |> 
  group_by(year) |> 
  summarize(total_cases = sum(cases))


# Visualize changes over time
ggplot(table1, aes(x = year, y = cases)) +
  geom_line(aes(group = country), color = "grey50") +
  geom_point(aes(color = country, shape = country)) +
  scale_x_continuous(breaks = c(1999, 2000)) # x-axis breaks at 1999 and 2000
```
###5.2.1 Exercises
For each of the sample tables, describe what each observation and each column represents.

#For table1, each observation is TB cases that happen in different countries and years among certain populations
#For table2, cases and population columns were combined into one column called type, each observation is reporting count of each type in different countries and years
#For table3, each observation is cases per population count(rate) in different countries and years

Sketch out the process you’d use to calculate the rate for table2 and table3. You will need to perform four operations:

1.  Extract the number of TB cases per country per year.
2.  Extract the matching population per country per year.
3.  Divide cases by population, and multiply by 10000.
4.  Store back in the appropriate place.

You haven’t yet learned all the functions you’d need to actually perform these operations, but you should still be able to think through the transformations you’d need.


```{r}
#1.  Extract the number of TB cases per country per year.
TB_cases <- table1 |> 
              select(country, year, cases)
TB_cases
#2.  Extract the matching population per country per year.
TB_population <- table1 |> 
              select(country, year, population)
TB_population




```


```{r}
#3.  Divide cases by population, and multiply by 10000.
rate_table <- table1 |> 
                mutate(rate = (cases / population) * 10000)
#4.  Store back in the appropriate place.
rate_table



```